Simple repression draft

### 2020-05-18 doc/paper.tex
Added more text on constitutive inference, plus redid bibfile

This is not wrong, but it is not the full story. The model's posterior is tightly constrained, but is it a good generative model? In other words, does the model generate data that ``look like'' our actual data, and is it therefore plausible that the model captures the important features of the data generating process? This intuitive notion can be codified with \textit{posterior predictive checks}, or PPCs, and we will see that this simple Poisson model fails badly.
\begin{figure}%[h!]
\centering
\includegraphics[width=0.9\textwidth]{../figures/fig2/fig2pt1.pdf}
\caption{\textbf{Constitutive promoter posterior
        inference and predictive checks.}
        (A) shows the joint posterior density from MCMC sampling of model 5,
        bursty promoter with negative binomially-distributed steady state.
        (B) shows the ECDF of the observed population distribution
        of mRNA transcripts under the control of a constitutive lacUV5 promoter.
        Posterior predictive percentiles are shown as colored bands for
        model (1), Poisson, and model (5), negative binomial.
        }
\label{fig:uv5_post}
\end{figure}

\mmnote{Outline:
\begin{itemize}
\item Poisson model. Write Bayes, briefly explain conjugacy and it's
convenience. immediately write down posterior, make Gaussian
approx. Explain PPC sampling (actually PP dist is analytically
known but you still have to draw RNGs from it. Also this obscures
the intuitive meaning of PPC sampling from posterior and
generating synthetic data, so I think it's more useful to just
generate PP samples from posterior draws, this gives a more
intuitive apples-to-apples comparison w/ model 5).
\item Bursty model. Write Bayes, sketch story of NB likelihood, sketch choice
of priors, show~\fig{fig:uv5_post}, see bursty $\gg$ Poisson model.
\item All 18 promoters. Hey the model still works on things other than
UV5, cool. Interesting and somewhat counterintuitive scaling b/w
burst rate and binding energy. Puzzle in comparing w/ Chong2014
supercoiling model. Are we missing important things?? Unclear, we
leave as open question. Good enough for our purposes: posterior is identifiable, PPC is great, and of the models we've thought of it is unique in satisfying both.
\end{itemize}
}

@@

% \bibliographystyle{nature}
\bibliographystyle{abbrv}
\bibliography{./library}

>>>>

This is not wrong, but it is not the full story. The model's
posterior is tightly constrained, but is it a good generative
model? In other words, does the model generate data that look similar
to our actual data, and is it therefore plausible that the
model captures the important features of the data generating
process? This intuitive notion can be codified with
\textit{posterior predictive checks}, or PPCs, and we will see
that this simple Poisson model fails badly.

The intuitive idea of posterior predictive checks is simple:
\mmnote{Not sure if all this stuff is clearer by example,
but then does it confuse the reader in generalizing to other models?}
\begin{enumerate}
\item Make a random draw of the model parameter $\lambda$
from the posterior distribution.
\item Plug that draw into the likelihood and generate
a synthetic dataset $\{m_k\}$ conditioned on $\lambda$.
\item Repeat many times.
\end{enumerate}
More formally, the posterior predictive distribution can be thought of
as the distribution of future yet-to-be-observed data,
conditioned on the data we have already observed.
Clearly if those data appear quite different,
the model has a problem. Put another way, if we suppose
the generative model is true, then the synthetic datasets
we generate should resemble the actual observed data, and if not,
it suggests the model is missing important features.
All the data we consider in this work are 1D (distributions of mRNA counts
over a population) so ECDFs are an excellent visual means of comparing
synthetic and observed datasets. In general for higher dimensional datasets,
much of the challenge is in merely designed good visualizations
that can actually show if synthetic and observed data are similar or not.

For our example Poisson promoter model then, we merely draw many
random numbers, say 1000, from the Gaussian posterior. For each one
of those draws, we generate a dataset from the likelihood,
i.e., we draw 2648 (the number of observed cells in the actual dataset)
Poisson-distributed numbers for each of the 1000 posterior draws,
for a total of 2648000 samples from the posterior predictive distribution.

To compare so many samples with the actual observed data,
one excellent visualization for 1D data is ECDFs of the quantiles,
as shown for our Poisson model in~\fig{fig:uv5_post}.
In this example, the median for each possible mRNA count is shown
as a dark green line, while shaded bands of progressively lighter green
contain 25\%, 50\%, 75\%, and 95\% of the posterior predictive samples.
Plotting quantiles in this way gives us a sense of the range of data
we might consider plausible, under the assumption that the model is true.
In this case it is quite obvious that the observed data, plotted in orange,
could not plausibly come from this Poisson generative model.
The other model with a PPC plotted in~\fig{fig:uv5_post},
bursty model 5 with a negative binomial steady-state mRNA distribution,
looks like a good candidate by eye. We cover it in more detail later.

\begin{figure}%[h!]
\centering
\includegraphics[width=0.9\textwidth]{../figures/fig2/fig2pt1.pdf}
\caption{\textbf{Constitutive promoter posterior
        inference and predictive checks.}
        (A) shows the joint posterior density from MCMC sampling of model 5,
        bursty promoter with negative binomially-distributed steady state.
        (B) shows the ECDF of the observed population distribution
        of mRNA transcripts under the control of a constitutive lacUV5 promoter.
        Posterior predictive percentiles are shown as colored bands for
        model (1), Poisson, and model (5), negative binomial.
        }
\label{fig:uv5_post}
\end{figure}

\mmnote{Outline, still to cover:
\begin{itemize}
\item Bursty model. Write Bayes, sketch story of NB likelihood, sketch choice
of priors, show~\fig{fig:uv5_post}, see bursty $\gg$ Poisson model.
\item All 18 promoters. Hey the model still works on things other than
UV5, cool. Interesting and somewhat counterintuitive scaling b/w
burst rate and binding energy. Puzzle in comparing w/ Chong2014
supercoiling model. Are we missing important things?? Unclear, we
leave as open question. Good enough for our purposes: posterior is identifiable, PPC is great, and of the models we've thought of it is unique in satisfying both.
\end{itemize}
}

@@

% \bibliographystyle{nature}
\bibliographystyle{abbrv}
\bibliography{./simple_rep}
